They are Very Smart: Genetic-Algorithm-Based Pathfinding in Games

### 2018-12-12 paper.tex
Paper: ALL THE TEXTTT

The GenomeManager class keeps track of game-wide genome management, i.e. selection and reproduction, as well as spawning Zombunnies and initiating new waves. The GenomeManager holds onto a collection of all genomes in a wave-by-wave list of hashmaps that map genomes to their fitness. The primary properties of the GenomeManager class can be seen in Figure 9, with various hardcoded constants that can be adjusted to influence general properties of the evolution, as well as the genomeMaps object, which contains the aforementioned collection of all genomes used in the course of evolution thus far.

@@

The initialization process of GenomeManager (see Figure 10) consists primarily of allocating the containing data structures for the various stages of the genome life cycle, followed by creating a first wave with randomized genomes and zero fitness. It should also be noted that we initialize our ZigguratGaussianSampler here from the Redzen library, which provides a fast way of generating normally-distributed floating-point numbers.

@@

The Spawn method, as seen in Figure 12, is called at every SPAWN\_DELAY interval POPULATION\_COUNT times. The function starts by spawning a Zombunny with no genome due to the inability to provide parameters upon GameObject instantiation. Therefore, we set the genome after the fact, popping it off of the toSpawn stack, also adding the ZombieDied function to the removed event on said Zombunny, which allows us to execute code in the context of the GenomeManager upon Zombunny death; namely, ZombunnyDied. ZombunnyDied handles finding the fitness of Zombunnies and adding them to the list of completed genomes, as well as calling NextWave again when the last Zombunny in the wave dies.

@@

ZombieAgent is a GameObject which determines Zombunny behavior and stats from a genome and keeps track of fitness for each individual Zombunny. To do so, ZombieAgent pulls all the stat-determining values from the genome and applies them upon instantiation, i.e. setting health, damage, and speed as determined by the genome, as well as handling tick-to-tick operations during a Zombunny's lifetime. These tick-to-tick operations include choosing the next movement vector based on the location of nearby GameObjects and agents' current status.

@@

In Figure 13, observe that the ZombieAgent class inherits from the Targetable class. Targetable is the parent class for every GameObject in the Action Game Framework, which can be automatically fired upon by ranged weapons, i.e. targeted. This also allows any existing defenses from the default game or any other targeting GameObject from the Action Game Framework to work without extra compatibility code. SINK\_SPEED and SINK\_TIME are constants used for death animation. velocity provides the Zombunny's current velocity for any consumers of the Targetable parent class. LastVelocity stores the previous movement vector as a PolarVector, in a format more congenial to the pathfinding algorithm. Fitness is an interface used to grab the effectiveness of this ZombieAgent at the end of its lifespan by the GenomeManager class. inVision keeps track of all nearby objects that the Zombunny can see for use in its pathfinding algorithm. the alignment is used for purposes of checking whether a Zombunny can damage other GameObjects, while visionCollider and damageCollider are the areas within which a Zombunny can see and deal damage from, respectively. The Rigidbody rigidBody handles physical collisions for movement purposes, while anim keeps track of the Zombunny's animation state, namely walking, idling, or death.

@@

In Figure 14, one can see the Genome, Awake, and Start methods of the
ZombieAgent class. Genome instantiates a Zombunny's Genome object by pulling the
var health from the given Genome, calling the SetMaxHealth and SetHealth
methods, and setting the initial velocity vector to have an initial direction of
$270^\circ$\----that is, in the general direction of the player's base relative
to the spawn location, and magnitude determined by the MovementSpeed variable
from the given Genome. Awake calls the underlying (base) Awake function and adds
custom OnDeath behavior to a Zombunny. Start initializes the inVision HashSet of
Targetable objects, which keeps track of nearby visible GameObjects. In the
unexpected case that genome or lastVelocity are null, Start initializes these to
default values (the "Zero" genome and a default velocity vector).

\imfigH{ZombieAgentVision}{Methods for keeping track of GameObjects entering
  and exiting a Zombunny's visionCollider}

Figure 15 shows the OnTriggerEnter and OnTriggerExit methods of the ZombieAgent
class, which are used in conjunction with a Zombunny's visionCollider to update
the inVision HashSet whenever a GameObject enters or exits the Zombunny's
visionCollider radius.

\imfigH{ZombieAgentUpdateNearby}{The portion of the FixedUpdate method for
determining the position of nearby objects relative to a Zombunny}

Figure 16 shows the functions used to determine the position of GameObjects in
inVision relative to a Zombunny's previous direction vector, stored as a
dictionary of Targetables to PolarVectors. This allows a Zombunny to keep track
of visible GameObjects relative to where it is currently facing.

\imfig{ZombieAgentUpdateMovement}{The function of FixedUpdate that handles
  movement between ticks}

Figure 17 shows the portion of FixedUpdate dedicated to altering a Zombunny's
trajectory from tick to tick, calling the Rotate method of the lastVelocity
PolarVector using the CalculateDirection method of the genome class, taking into
account the Zombunny's previous velocity vector, time alive, current health, and
the aforementioned dictionary of nearby GameObjects.

\imfig{ZombieAgentUpdateDeath}{The portion of FixedUpdate that handles
  death/deletion}

Figure 18 shows the part of the FixedUpdate method which keeps track of Zombunny
death, destroying the Zombunny GameObject at the appropriate time and calling
the relevant death animations. For every tick that a Zombunny is not dead,
FixedUpdate increments timeAlive, which is later used as a factor in
calculating fitness.

@@

The ZombieAttacker class (see Figure 19) calls Update on each tick, dealing damage to any item
in toAttack, which is a HashSet of Targetable GameObjects that intersect with
a Zombunny's attack collider (see Figure 4). This way, a Zombunny can attack
multiple GameObjects so long as they intersect with its attack collider. Damage
is applied to items in toAttack by calling their individual takeDamage methods.
Damage dealt is determined by the Zombunny's genome, and is passed as a parameter
to takeDamage.

\imfig{ZombieAttacker}{The Update method of the ZombieAttacker class}

>>>>

The \texttt{GenomeManager} class keeps track of game-wide state and genome tasks, i.e.
selection and reproduction, as well as spawning Zombunnies and initiating new
waves. The \texttt{GenomeManager} holds onto a collection of all genomes in a
wave-by-wave list of hashmaps that map genomes to their fitness. The primary
properties of the \texttt{GenomeManager} class can be seen in Figure 9, with various
hardcoded constants that can be adjusted to influence general properties of the
evolution, as well as the genomeMaps object, which contains the aforementioned
collection of all genomes used in the course of evolution thus far.

@@

The initialization process of \texttt{GenomeManager} (see Figure 10) consists primarily of allocating the containing data structures for the various stages of the genome life cycle, followed by creating a first wave with randomized genomes and zero fitness. It should also be noted that we initialize our \texttt{ZigguratGaussianSampler} here from the Redzen library, which provides a fast way of generating normally-distributed floating-point numbers.

@@

The \texttt{Spawn} method, as seen in Figure 12, is called at every
\texttt{SPAWN\_DELAY} interval \texttt{POPULATION\_COUNT} times. The function
starts by spawning a Zombunny with no genome due to the inability to provide
parameters upon GameObject instantiation. Therefore, we set the genome after the
fact, popping it off of the \texttt{toSpawn} stack, also adding the \texttt{ZombieDied} function
to the removed event on said Zombunny, which allows us to execute code in the
context of the \texttt{GenomeManager} upon Zombunny death; namely, \texttt{ZombunnyDied}.
ZombunnyDied handles finding the fitness of Zombunnies and adding them to the
list of completed genomes, as well as calling \texttt{NextWave} again when the last
Zombunny in the wave dies.

@@

\texttt{ZombieAgent} is a \texttt{GameObject} which determines Zombunny behavior
and stats from a genome and keeps track of fitness for each individual Zombunny.
To do so, \texttt{ZombieAgent} pulls all the stat-determining values from the
genome and applies them upon instantiation, i.e. setting health, damage, and
speed as determined by the genome, as well as handling tick-to-tick operations
during a Zombunny's lifetime. These tick-to-tick operations include choosing the
next movement vector based on the location of nearby \texttt{Targetables} and agents' current status.

@@

In Figure 13, observe that the \texttt{ZombieAgent} class inherits from the
\texttt{Targetable} class. \texttt{Targetable} is the parent class for every
GameObject in the Action Game Framework that can be automatically
fired upon by ranged weapons, i.e. targeted. This also allows any existing
defenses from the default game or any other targeting GameObject from the Action
Game Framework to work without extra compatibility code. \texttt{SINK\_SPEED} and
\texttt{SINK\_TIME} are constants used for death animation. \texttt{velocity} provides the
Zombunny's current velocity for any consumers of the \texttt{Targetable} parent class.
\texttt{LastVelocity} stores the previous movement vector as a \texttt{PolarVector}, in a format
more congenial to the pathfinding algorithm. \texttt{Fitness} is an interface used to
grab the effectiveness of this \texttt{ZombieAgent} at the end of its lifespan by the
\texttt{GenomeManager} class. \texttt{inVision} keeps track of all nearby objects that the
Zombunny can see for use in its pathfinding algorithm. \texttt{alignment} is used for
purposes of checking whether a Zombunny can damage other \texttt{Targetables}, while
\texttt{visionCollider} and \texttt{damageCollider} are the areas within which a Zombunny can see
and deal damage from, respectively. The Rigidbody \texttt{rigidBody} handles physical
collisions for movement purposes, while \texttt{anim} keeps track of the Zombunny's
animation state, namely walking, idling, or death.

@@

In Figure 14, one can see the \texttt{Genome}, \texttt{Awake}, and \texttt{Start} methods of the
\texttt{ZombieAgent} class. \texttt{Genome} instantiates a Zombunny's Genome object by pulling the
health from the given \texttt{Genome,} calling the \texttt{SetMaxHealth} and \texttt{SetHealth}
methods, and setting the initial velocity vector to have an initial direction of
$270^\circ$\----that is, in the general direction of the player's base relative
to the spawn location, and magnitude determined by the \texttt{MovementSpeed} variable
from the given \texttt{Genome}. \texttt{Awake} calls the underlying (base) \texttt{Awake} function and adds
custom \texttt{OnDeath} behavior to a Zombunny. \texttt{Start} initializes the \texttt{inVision} \texttt{HashSet} of
\texttt{Targetable} objects, which keeps track of nearby visible \texttt{GameObjects}. In the
unexpected (this is a concurrency issue) case that \texttt{genome} or \texttt{lastVelocity} are \texttt{null}, \texttt{Start} initializes these to
default values (the "Zero" genome and a default velocity vector).

\imfigH{ZombieAgentVision}{Methods for keeping track of \texttt{GameObjects} entering
  and exiting a Zombunny's \texttt{visionCollider}}

Figure 15 shows the \texttt{OnTriggerEnter} and \texttt{OnTriggerExit} methods of the \texttt{ZombieAgent}
class, which are used in conjunction with a Zombunny's \texttt{visionCollider} to update
the \texttt{inVision} \texttt{HashSet} whenever a \texttt{GameObject} enters or exits the Zombunny's
\texttt{visionCollider} radius.

\imfigH{ZombieAgentUpdateNearby}{The portion of the \texttt{FixedUpdate} method for
determining the position of nearby objects relative to a Zombunny}

Figure 16 shows the functions used to determine the position of GameObjects in
\texttt{inVision} relative to a Zombunny's previous direction vector, stored as a
dictionary of Targetables to PolarVectors. This allows a Zombunny to keep track
of visible GameObjects relative to where it is currently facing.

\imfig{ZombieAgentUpdateMovement}{The function of \texttt{FixedUpdate} that handles
  movement between ticks}

Figure 17 shows the portion of \texttt{FixedUpdate} dedicated to altering a Zombunny's
trajectory from tick to tick, calling the \texttt{Rotate} method of the \texttt{lastVelocity}
\texttt{PolarVector} using the \texttt{CalculateDirection} method of the genome class, taking into
account the Zombunny's previous velocity vector, time alive, current health, and
the aforementioned dictionary of nearby GameObjects.

\imfig{ZombieAgentUpdateDeath}{The portion of \texttt{FixedUpdate} that handles
  death/deletion}

Figure 18 shows the part of the \texttt{FixedUpdate} method which keeps track of Zombunny
death, destroying the Zombunny GameObject at the appropriate time and calling
the relevant death animations. For every tick that a Zombunny is not dead,
\texttt{FixedUpdate} increments \texttt{timeAlive}, which is later used as a factor in
calculating fitness.

@@

The \texttt{ZombieAttacker} class (see Figure 19) calls \texttt{Update} on each tick, dealing damage to any item
in \texttt{toAttack}, which is a \texttt{HashSet} of \texttt{Targetable} GameObjects that intersect with
a Zombunny's attack collider (see Figure 4). This way, a Zombunny can attack
multiple GameObjects so long as they intersect with its (tiny) attack collider.
Ideally the Zombunny would only attack one other entitiy at a time, however
attempting to do so results in a lot of edge cases and extra computational work,
so we opted to attack everything in range. Damage
is applied to items in \texttt{toAttack} by calling their individual
\texttt{takeDamage} methods. The amount of damage dealt is determined by the Zombunny's genome, and is passed as a parameter
to \texttt{takeDamage}.

\imfig{ZombieAttacker}{The \texttt{Update} method of the \texttt{ZombieAttacker} class}
